### 2015-12-07 paper/4-performance.tex
Review

%TODO say that you will now prove your claims about the aproach having nearly no impact
%TODO and potentially even claim that you are faster than adevs for some situations, due to the algorithms from PythonPDEVS and further features
%TODO briefly give some hardware info on the machine (CPU, OS, RAM), used software (version of adevs, GCC, ...), and about number of samples/variance
\subsection{Sequential Simulation}
%TODO we start with sequential simulation results, because this is still base case
%TODO compare CPU time with adevs for some benchmarks
%TODO and explain why there is such a difference
%TODO (spend some time here, as this is also important)
\subsubsection{Preamble}
The benchmark models are identical in functionality (between adevs/dxex versions). Dxex uses integer timestamps, adevs uses floating point. Tests show there is little difference performance-wise (both types fit in a machine word).\\ The benchmarks are compiled using O3 and link-time optimization, with all IO disabled. The benchmarks have testing variants where the exact output can be verified.
Benchmarks ran on a i7-2670QM, 16GiB Ram, Arch Linux (4.2.5 kernel), G++ 5.2. Interpretation of the benchmark results is done using call-graphs generated by perf, a selection of those call-graphs can be found in the code repository. For the runtime plots perf is used in combination with R.\\
Note that a fixed timeadvance (e.g. devstone) is set at 100.
\subsubsection{Devstone}
The Devstone \cite{DEVStone} benchmark is highly hierarchical, using Direct connection the dxex kernels can exploit this whereas adevs needs to walk the coupled model's structure to pass events.\\
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig1.eps}
	\label{fig1.eps}
\end{figure}

\subsubsection{PHold}
PHold \cite{PHOLD} is a parallel oriented benchmark, sequential runtime is measured only as a baseline. A model selects the destination of a message at runtime using a state-saved RNG, which takes up a non-trivial amount of runtime.\\
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig4.eps}
	\label{fig4.eps}
\end{figure}

\subsubsection{Interconnect}
Interconnect \cite{van2013research} is a benchmark where all models broadcast, creating a complete graph in dependencies between models. As the model count increases, we see the expected quadratic increase in runtime for both adevs and dxex, but an increasing penalty for dxex. Profiling shows this is entirely due to the heap allocation of messages, which even though minimized by using memory pools remains significant.\\
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig3.eps}
	\label{fig3.eps}
\end{figure}

\subsection{Parallel Simulation}
%TODO explain that we will now compare different synchronization protocols
%TODO compare different synchronization protocols with adevs' conservative, showing that we are faster (or at least, not that much slower) than adevs
Whenever not specified, the benchmarks use 4 kernels for parallel simulation.
\subsubsection{Devstone}
The flattened models are allocated to kernels by giving each kernel a distinct section of the chain, resulting in a low ratio of inter-kernel to intra-kernel messages. For optimistic, this can cause more reverts since the kernels will start to drift faster as the model-count increases. Furthermore, optimistic is quite sensitive to an increase in kernels, since the delay before a revert propagates increases. Of note as well is the warm-up time this benchmark requires, for $n=d\times w$ models, it takes $\textup{timeadvance}()*(n-1)$ transitions to activate the last model in the chain. For parallel this can be reduced to $\textup{timeadvance}()*(n\frac{kernels-1}{kernels} -1)$ before the last kernel becomes active.\\
\begin{figure}[ltbh]
	\includegraphics[width=.5\textwidth]{fig/fig2.eps}
	\label{fig2.eps}
\end{figure}
\subsubsection{PHold}
%Phold
In Phold allocation is specified in the benchmark itself. Each kernel manages a single node with a constant set of sub-nodes. The parameter R (here 10) determines the percentage of remote destination models.\\
The dynamic dependency graph is a very sparse version of the static dependency graph, penalizing conservative. Lookahead is $\epsilon$, so conservative spends most of its time crawling in steps of $\epsilon$. Since the dependency graph between kernels is a complete graph, this is not a simulation that scales in our implementation. For N kernels, each kernel has to query the null-time of N-1 kernels, resulting in O($N^2$) polling behaviour. In a non-cyclic simulation with a non-trivial lookahead (e.g. devstone), that choice however does pay off. Adevs's lesser performance is due in part to their lookahead management, which after profiling shows to spend a non-trivial amount of time in exception handling code. 
Optimistic suffers little from the above problems, however due to the high interconnectivity a cascading revert is still possible. With the percentage of remotes = 100 phold reflects interconnect in behaviour, which is why the R value is not dimensioned here. A revert is very expensive in PHold due to our usage of C++11's random nr generators. The cost of a revert is dominated by the recalculation of destination models, not in allocating/deallocating states/messages. Once a revert happens the drift between kernels increases fast, increasing the likelihood of more reverts. Despite all this, optimistic can for low R values quickly exploit the uncertainty that slows down conservative in this benchmark.

\subsubsection{Interconnect}
In Interconnect
the set of atomic models form a complete graph (w.r.t. connections), each model broadcasts messages to the entire set. \\
Allocation cannot avoid cycles, the resulting dependency graph between kernels remains a complete graph. The runtime dependency graph is almost immediately equal to the static graph.
Conservative still faces the same issues as in PHold, with the key difference that for a fixed time advance lookahead is equal to the timespan between transitions. The scaling issue is identical as with PHold.\\
In this benchmark optimistic runs very quickly out of memory. With c kernels and N atomic models, a single revert undoing k transitions will lead to $(N-1)\times k$ messages that need to be recreated, plus $(N-1) \frac{c-1}{c} \times k$ anti-messages that need to be send.

\subsubsection{Priority network model}
%TODO show a single plot which varies a parameter that changes the ideal synchronization protocol
The priority benchmark is composed of a single server generating a stream of $0<=m<=n$ messages at fixed time intervals, interleaved with a probability p for a priority message, to n receivers. \\ This defaults lookahead for the receivers to $\epsilon$ but this time there is no scaling effect, nor are there cycles in the dependency graph. This model therefore highlights the basic strengths/weaknesses of both synchronization protocols. Receiving models are allocated on another kernel than the server, and have a internal transition so that they will not wait for the incoming messages. 
\input pqueue.tex \\
A key difference here with the other benchmarks is that a state (in the Receiver instances) is very cheap to copy/create. The kernel holding the server will never revert since it is a source in the dependency graph. Optimistic will therefore not suffer the same performance hit in recreating states as it does in PHold. The overhead in optimistic is entirely due to the factor m, which will quickly dominate in increasing buffers of received messages.
\begin{figure}[ltbh]
	\includegraphics[width=.5\textwidth]{fig/fig5.eps}
	\label{fig5.eps}
\end{figure}
Interestingly, the parameter p does not clearly benefit either synchronization protocol. While this removes any possibility for a lookahead, conservative can quickly bridge the timespan between fixed messages since there is no cyclic dependency. Lightweight states prevent performance loss due to reverts in optimistic, only the overhead in event handling in optimistic eventually becomes the deciding factor.

@@

Both dxex and adevs use tcmalloc as memory allocator. Additionally, dxex uses memory pools to further reduce the frequency of expensive system calls (malloc/free/sbrk/mmap/...). Tcmalloc will only gradually release memory back to the OS, whereas our pools will not do so at all. If memory has been allocated once, it is from a performance point of view better to keep that memory in the pool. This is one reason why memory utilization is best measured by peak allocation. Profiling is done using Valgrind's massif tool. \cite{Nethercote:2007:VFH:1273442.1250746}
Platform used for memory profiling has an i5-3317U Intel cpu and 8GiB RAM with a page size of 4,096KiB, running Fedora 22 (kernel 4.2.6).\\

@@

Adevs passes messages by value, we pass by pointer. The runtime effects of this choice are already demonstrated in the interconnect benchmark, so in this section, we measure memory usage in number of allocated pages combining text, stack and heap memory for the program profiled. For the OS and/or user this is the actual memory footprint of the application. It is important to note that, especially in the case of optimistic, not all this memory is always in use by the kernels. During simulation, the pools will generally not return deallocated memory to the OS, but keep it for later reuse.

@@

@@

@@

		Adevs & AdevsCon &DX &DXCon &DXOpt\\ \hline
Since conservative passes messages by pointer, it needs a GVT/LBTS implementation to organise garbage collection. This inevitable delay explains the higher memory usage compared to adevs.\\
Optimistic's TimeWarp requires state/event saving, and its GVT algorithm is more complex (with a resulting higher latency) than the LBTS calculation in conservative. 
Additionally, the differences in LP virtual times are far larger compared to conservative time synchronization. All these factors explain the heavier memory usage. Devstone (flattened) is allocated in a chain. Leafs in the dependency graph will therefore do a lot of unnecessary simulation before having a revert, leading to increased memory pressure. Unlike conservative and sequential execution, memory usage in optimistic varies greatly depending on scheduling of kernel threads and drifting between kernels. 
		Adevs & AdevsCon &DX &DXCon &DXOpt\\ \hline

@@

		Adevs & AdevsCon &DX &DXCon & DXOpt\\ \hline

@@

\paragraph*{Priority Network}
The priority network model is detailed in section 4.1.\\
Conservative's high memory usage is largely due to the garbage collection implementation. Kernel 0 is the initiator of the lbts calculation but since it holds only models independent of any other (server), it will finish simulation very fast leaving the other kernel bereft of an updated lbts. The kernel holding the server will after simulation wait on the receiving kernel until it can prove all sent messages can be deallocated.
\begin{table}[lhtb]
	\centering
	\caption{Priority model n 128, m 16, p 10,  t2e8, unit MiB, 2 kernels}
	\label{pmod_mem}
	\begin{tabular}{| l | l | l |}
		\hline
		DX &DXCon &DXOpt\\ \hline
		35 & 450 & 651\\ \hline
	\end{tabular}
\end{table}

>>>>

%TODO say that you will now prove your claims about the aproach having nearly no impact
%TODO and potentially even claim that you are faster than adevs for some situations, due to the algorithms from PythonPDEVS and further features
%TODO briefly give some hardware info on the machine (CPU, OS, RAM), used software (version of adevs, GCC, ...), and about number of samples/variance
\subsection{Sequential Simulation}
%TODO we start with sequential simulation results, because this is still base case
%TODO compare CPU time with adevs for some benchmarks
%TODO and explain why there is such a difference
%TODO (spend some time here, as this is also important)
In our benchmarks, we will consistently compare the performance of dxex with adevs. The implemented models for the each benchmark are functionally identical.\\
Dxex uses integer timestamps whereas adevs uses floating point time representation. Tests show there is little difference performance-wise (both types fit in a machine word).\\
The benchmarks have been compiled using the {\tt -O3} option and link-time optimization, with all IO disabled. To compare the models in details, we have run the different benchmarks with different parameters.\\
The benchmarks have been run on a i7-2670QM, 16GiB RAM, Arch Linux (4.2.5 kernel) and G++ 5.2. Interpretation of the benchmark results is done using call-graphs generated by perf. A selection of those call-graphs is available in the code repository. For the runtime plots, perf is used in combination with R.\\
%Note that a fixed timeadvance (e.g. devstone) is set at 100.
\subsubsection{Devstone}
The Devstone \cite{DEVStone} benchmark is highly hierarchical. As can be seen in Figure~\ref{fig:Devstone}, dxex is outperforming adevs considerably. This can be explained by the fact that, using Direct connection, the hierharchical structure is being flattened whereas, in the adevs case, considerable overhead is caused by traversing the coupled model's structure to pass events.
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig1.eps}
	\caption{Devstone}
	\label{fig:Devstone}
\end{figure}

\subsubsection{PHold}
PHold \cite{PHOLD} is a benchmark that focuses in particular on parallel execution; the sequential runtime is therefore only added as a baseline. A model selects the destination of a message at runtime using a state-saved RNG, which takes up a non-trivial amount of runtime. The results of the benchmark are visualized in Figure~\ref{fig:PHold}.
%% needs further clarification...
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig4.eps}
	\caption{PHold}
	\label{fig:PHold}
\end{figure}

\subsubsection{Interconnect}
Interconnect~\cite{van2013research} is a benchmark where all models broadcast, creating a complete graph in dependencies between models. As the model count increases, we see (see Figure~\ref{fig:Interconnect} the expected quadratic increase in runtime for both adevs and dxex, but an increasing penalty for dxex. Profiling shows this is entirely due to the heap allocation of messages, which even though minimized by using memory pools remains significant.
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig3.eps}
	\label{fig3.eps}
	\caption{Interconnect}
	\label{fig:Interconnect}
\end{figure}

\subsection{Parallel Simulation}
%TODO explain that we will now compare different synchronization protocols
%TODO compare different synchronization protocols with adevs' conservative, showing that we are faster (or at least, not that much slower) than adevs
By default, the benchmarks use 4 kernels for parallel simulation.
\subsubsection{Devstone}
The flattened models are allocated to kernels by giving each kernel a distinct section of the chain, resulting in a low ratio of inter-kernel to intra-kernel messages. For the optimistic case, this can cause more reverts since the kernels will start to drift faster as the model-count increases. Furthermore, optimistic synchronization is quite sensitive to an increase in kernels, since the delay before a revert propagates, increases. This benchmark requires a specific warm-up time; for $n=d\times w$ models, it takes $\textup{timeadvance}()*(n-1)$ transitions to activate the last model in the chain. When using parallel processing, this can be reduced to $\textup{timeadvance}()*(n\frac{kernels-1}{kernels} -1)$ before the last kernel becomes active.\\
Figure~\cite{fig:DevstonParallel} shows that dxex outperformance adevs (in both optimistic and conservative case).
\begin{figure}[ltbh]
	\includegraphics[width=.5\textwidth]{fig/fig2.eps}
	\label{fig:DevstoneParallel}
	\caption{DevStone parallel}
\end{figure}
\subsubsection{PHold}
%Phold
In Phold, the allocation is specified in the benchmark itself. Each kernel manages a single node with a constant set of sub-nodes. The parameter $R$ (which is here set to $10$) determines the percentage of remote destination models.\\
The dynamic dependency graph is a very sparse version of the static dependency graph, penalizing the conservative case. The lookahead is $\epsilon$, so the conservative case spends most of its time crawling in steps of $\epsilon$. Since the dependency graph between kernels is a complete graph, this is not a simulation that scales in our implementation. For $N$ kernels, each kernel has to query the null-time of $N-1$ kernels, resulting in O($N^2$) polling behaviour. In a non-cyclic simulation with a non-trivial lookahead (like in, e.g., Devstone), that choice does pay off (see Figure~\ref{fig:DevstoneParallel}). Adevs's lesser performance is due in part to their lookahead management, which after profiling shows to spend a non-trivial amount of time in exception handling code.\\
The optimistic case suffers little from the above problems; due to the high interconnectivity, however, a cascading revert is still possible. With the percentage of remotes equal to $100$, Phold reflects interconnect in behaviour, which is why the $R$ value is not dimensioned here. A revert is very expensive in PHold due to our usage of C++11's random nr generators. The cost of a revert is dominated by the recalculation of destination models, not in allocating/deallocating states/messages. Once a revert happens the drift between kernels increases fast, increasing the likelihood of more reverts. Despite all this, optimistic can for low $R$ values quickly exploit the uncertainty that slows down conservative in this benchmark.

\subsubsection{Interconnect}
In Interconnect, the set of atomic models form a complete graph (w.r.t. connections); each model broadcasts messages to the entire set.\\
Allocation cannot avoid cycles and the resulting dependency graph between kernels remains a complete graph. The runtime dependency graph is almost immediately identical to the static graph.\\
The conservative case still shows the same issues as in PHold, with the key difference , for a fixed time advance, the lookahead is equal to the timespan between transitions. The scaling issue is identical as with PHold.\\
In this benchmark, the optimistic case runs very quickly out of memory. With $c$ kernels and $N$ atomic models, a single revert undoing $k$ transitions will lead to $(N-1)\times k$ messages that need to be recreated, plus $(N-1) \frac{c-1}{c} \times k$ anti-messages that need to be sent.

\subsubsection{Priority network model}
%TODO show a single plot which varies a parameter that changes the ideal synchronization protocol
The priority benchmark is composed of a single server generating a stream of $0\leq m \leq n$ messages at fixed time intervals, interleaved with a probability $p$ for a priority message to $n$ receivers (see Figure~\ref{fig:NetworkModel}).\\ This defaults the lookahead for the receivers to $\epsilon$, but this time there is no scaling effect, nor are there cycles in the dependency graph. This model therefore highlights the basic strengths/weaknesses of both synchronization protocols. Receiving models are allocated on another kernel than the server and have an internal transition so that they will not wait for the incoming messages.
\begin{center}
\begin{figure}
\input pqueue.tex
\caption{The priority network model}
\label{fig:NetworkModel}
\end{figure}
\end{center}
A key difference here with the other benchmarks is that a state (in the Receiver instances) is very cheap to copy/create. The kernel holding the server will never revert since it is a source in the dependency graph. The optimistic case will therefore not suffer the same performance hit in recreating the states as it does in PHold. The overhead in the optimistic case is entirely due to the factor $m$, which will quickly dominate in increasing buffers of received messages.
\begin{figure}[ltbh]
	\includegraphics[width=.5\textwidth]{fig/fig5.eps}
	\label{fig:Priority}
	\caption{Priority}
\end{figure}
Interestingly, the parameter $p$ does not clearly favour either synchronization protocol (see Figure~\ref{fig:Priority}). While this removes any possibility for a lookahead, the conservative case can quickly bridge the timespan between fixed messages since there is no cyclic dependency. Lightweight states prevent performance loss due to reverts in the optimistic case; only the overhead in event handling in the optimistic case eventually becomes the deciding factor.

@@

Both dxex and adevs use tcmalloc as memory allocator. Additionally, dxex uses memory pools to further reduce the frequency of expensive system calls (malloc/free/sbrk/mmap/...). Tcmalloc will only gradually release memory back to the OS, whereas our pools will not do so at all. If memory has been allocated once, it is from a performance point of view better to keep that memory in the pool. This is one reason why memory utilization is best measured by peak allocation. Profiling is done using Valgrind's massif tool~\cite{Nethercote:2007:VFH:1273442.1250746}.
The platform used for memory profiling has an i5-3317U Intel CPU and 8GiB RAM with a page size of 4,096KiB, running Fedora 22 (kernel 4.2.6).

@@

Adevs passes messages by value, dxex passes a pointer. The runtime effects of this choice have already been demonstrated in the Interconnect benchmark, so, in this section, we measure memory usage in number of allocated pages combining text, stack and heap memory for the program profiled. For the OS and/or user, this is the actual memory footprint of the application. It is important to note that, especially in the optimistic case, not all this memory is always in use by the kernels. During simulation, the pools will generally not return deallocated memory to the OS, but keep it for later reuse.

@@

		adevs & adevs con &dxex &dxex con&dxex opt\\ \hline

@@

Since, in the conservative case, messages are passed by pa ointer, a GVT/LBTS implementation is required to organise the garbage collection. This inevitable delay explains the higher memory usage compared to adevs.\\
Optimistic's TimeWarp requires state/event saving, and its GVT algorithm is more complex (with a resulting higher latency) than the LBTS calculation in the conservative case. 
Moreover, the differences in LP virtual times are far larger compared to conservative time synchronization. All these factors explain the heavier memory usage. Devstone (flattened) is allocated in a chain. Leafs in the dependency graph will therefore do a lot of unnecessary simulation before having a revert, leading to an increased memory pressure. Unlike conservative and sequential execution, memory usage in the optimistic case varies greatly depending on scheduling of kernel threads and drifting between kernels. 

@@

		adevs & adevs con &dxex &dxex con &dxex opt\\ \hline

@@

		adevs & adevs con &dxex &dxex con & dxex opt\\ \hline

@@

\paragraph*{Priority Network}
The priority network model is detailed in section 4.1.\\
The high memory usage of the conservative case is largely due to the garbage collection implementation. Kernel 0 is the initiator of the LBTS calculation but since it holds only models independent of any other (server), it will finish simulation very fast leaving the other kernel bereft of an updated LBTS. The kernel holding the server will after simulation wait on the receiving kernel until it can prove all sent messages can be deallocated.
\begin{table}[lhtb]
	\centering
	\label{pmod_mem}
	\begin{tabular}{| l | l | l |}
		\hline
		dxex &dxex con &dxex opt\\ \hline
		35 & 450 & 651\\ \hline
	\end{tabular}
	\caption{Priority model n 128, m 16, p 10,  t2e8, unit MiB, 2 kernels}
\end{table}

### 2015-12-07 paper/5-related-work.tex
Review

%TODO explain about other DEVS simulation kernels, and how they differ
%TODO explain about PythonPDEVS, the most related in terms of feature and "legacy" --> now in C++11 and way better performance
\subsection{PythonPDEVS}
Dxex is closely related to PyPDEVS in design and philosophy. PyPDEVS allows anyone who grasps the PDEVS formalisms to immediately simulate his/her model without having to consider the kernel implementation.A Python based implementation offers the advantage of very fast prototype/run/evaluate cycles, this can't be matched by a C++ simulator but it should be noted that once the kernels are compiled (shared libraries), the actual compilation time of the model is small enough to make prototyping possible.
Advanced features such as activity based relocation and the performance gains this results in, are still unique to PyPDEVS.
\subsection{Adevs}
%TODO explain about adevs, the most related in terms of performance --> we implement more synchronization options and offer better performance because ...
Adevs is still under active development, allowing for an exact comparison in performance and features. It remains a very fast simulation engine for the PDEVS formalism, but it lacks an optimistic synchronization implementation. %Tracing ?
By virtue of not flattening Coupled Models, adevs's performance degrades in increasingly hierarchical models. Dxex manages lookahead with more overhead than adevs, leading to a performance difference as the model count increases. Finally, adevs employs the full dependency graph, contrasted to dxex kernels which only observe 1-edge removed nodes. 
%checkme
\subsection{CD++}
%TODO explain about CD++, the most related in terms of synchronization --> we implement everything in a single program, with a non-fragmented code-base
%TODO make mention of Warped, which is the kernel used in CD++, and why we don't use it
Different projects on CD++ offer conservative (CCD++) as well as optimistic (PCD++) parallel simulation. In contrast to our single program, with a non-fragmented code-base, neither projects offer both synchronization protocols. CD++ relies on the WARPED kernel. It is a middleware that provides memory, event, file, time and communication scheduling. We did not use the WARPED kernel (nor the underlying MPI), dxex is designed specifically for a shared memory architecture and as such any middleware, while feature-rich, would have lead to unacceptable overhead.

>>>>

%TODO explain about other DEVS simulation kernels, and how they differ
%TODO explain about PythonPDEVS, the most related in terms of feature and "legacy" --> now in C++11 and way better performance
\subsection{PythonPDEVS}
Dxex is closely related to PyPDEVS in design and philosophy. PyPDEVS allows anyone who grasps the PDEVS formalisms to immediately simulate his/her model without having to consider the kernel implementation. A Python based implementation offers the advantage of very fast prototype/run/evaluate cycles, this can't be matched by a C++ simulator. It should be noted that, once the kernels have been compiled (in shared libraries), the actual compilation time of the model is small enough to make prototyping possible.
Advanced features such as activity based relocation and the performance gains this results in, are still unique to PyPDEVS.
\subsection{Adevs}
%TODO explain about adevs, the most related in terms of performance --> we implement more synchronization options and offer better performance because ...
Adevs is still under active development, allowing for an exact comparison in performance and features. It remains a very fast simulation engine for the PDEVS formalism, but it lacks an optimistic synchronization implementation. %Tracing ?
Due to the fact that adev does not flatten the Coupled Models, it's performance degrades significantly if the hierarchy becomes deeper. Dxex, on the other hand, manages lookahead with more overhead than adevs, leading to a performance difference as the model count increases. Finally, adevs employs the full dependency graph this in contrast with dxex kernels, which only observe 1-edge removed nodes. 
%checkme
\subsection{CD++}
%TODO explain about CD++, the most related in terms of synchronization --> we implement everything in a single program, with a non-fragmented code-base
%TODO make mention of Warped, which is the kernel used in CD++, and why we don't use it
Different projects on CD++ offer conservative (CCD++) as well as optimistic (PCD++) parallel simulation. In contrast to dxex,  neither projects offer both synchronization protocols. CD++ relies on the WARPED kernel. It is a middleware that provides memory, event, file, time and communication scheduling. We did not use the WARPED kernel (nor the underlying MPI) because dxex is designed specifically for a shared memory architecture and, as such, any middleware, however feature-rich, would have lead to an unacceptable overhead.

### 2015-12-07 paper/6-conclusions.tex
Review

%TODO conclude the paper, saying that our new simulation tool offers conservative and optimistic synchronization, resulting in good performance, which is even better than adevs
Both synchronization implementations offer good performance in differing simulations, in some simulations dxex outperforms adevs whereas in others we can still improve. 
The optimistic implementation uses an aggressive TimeWarp variant, this needs to be extended with lazy evaluation/cancellation to function in cyclic simulations. 
\subsection{Future work}
%TODO future work: more things that we might want to implement, and probably something about automatically switching between synchronization protocols at runtime (and reference my activity papers ;))
\subsubsection{Activity}
As shown in \cite{PythonPDEVS_ACTIMS} activity and allocation of models across kernels is a key aspect in achieving high performance in any parallel implementation. Allocating models so that there are no dependency cycles between their containing kernels is a first step, but not always possible. For optimistic one can use re-allocation to break (runtime dependency) cycles and/or perform load-balancing. If kernels are unevenly balanced they will begin to drift fast, causing increasingly more reverts. There is a limited framework implemented to track model activity for debugging purposes, this could be extended to enable the above strategies.
\subsubsection{Hybrid}
The optimistic implementation could use (null/eot/eit) from conservative to detect and/or reduce the cost of reverts without completely stalling on influencing kernels.
Conservative kernels could be extended with runtime information about influencing models, for example if one can guarantee a static dependency is not used for a fixed time-span, this can be removed for that period of (virtual) time. 
Ultimately the simulation could switch at runtime between protocols based on the information provided by activity tracking. This requires the above mentioned activity tracking framework, and could only be done if a GVT/LBTS time is agreed upon between the kernels. From the model point of view no changes are needed, although a non-trivial lookahead is obviously desired.

>>>>

\section{Future work}
%TODO future work: more things that we might want to implement, and probably something about automatically switching between synchronization protocols at runtime (and reference my activity papers ;))
\subsection{Memory}
Our optimistic implementation can benefit from a faster GVT algorithm such
 as used in~\cite{Fujimoto:1997:CGV:268403.268404}  or, more recently in~\cite{Bauer:2005:SND:1069810.1070159}. 
With access to the memory subsystem already in place (pools), the optimistic case could be constrained using per kernel quota, preventing the simulation from exhausting memory. While less sensitive to memory pressure, the same approach could improve the conservative case.
\subsection{Activity}
As shown in~\cite{PythonPDEVS_ACTIMS}, activity and allocation of models across kernels is a key aspect in achieving high performance in any parallel implementation. Allocating models so that there are no dependency cycles between their containing kernels is a first step, but not always possible. For the optimistic case, one can use re-allocation to break (runtime dependency) cycles and/or perform load-balancing. If kernels are unevenly balanced, they will begin to drift fast, causing increasingly more reverts. In dxex, we implement a limited framework to track model activity for debugging purposes; this could be extended to enable the above strategies.
\subsection{Hybrid}
The optimistic implementation could use (null/eot/eit) from the conservative case to detect and/or reduce the cost of reverts without completely stalling on influencing kernels.\\
The eot calculation in the conservative case can be enhanced by using the entire dependency graph between kernels (not models). The implementation could detect dependency cycles between kernels and, instead of waiting on dependent kernels, advance as a synced group with better scaling. LBTS calculation could leverage this as well to reduce memory usage, though this is not that stringent in the conservative case.
Ultimately, the simulation could switch at runtime between protocols based on the information provided by activity tracking. This requires the above mentioned framework, and can only be done at timepoints where a GVT/LBTS time is agreed upon between the kernels. From the model's perspective, no changes are needed, although a non-trivial lookahead is obviously desired.

\section{CONCLUSIONS}
%TODO conclude the paper, saying that our new simulation tool offers conservative and optimistic synchronization, resulting in good performance, which is even better than adevs
In this paper, we presented dxex, a new C++-based PDEVS implementation supporting both sequential and parallel processing models.\\
We showed that, in the sequential case, dxex outperforms adevs in hierarchical simulations and that adevs produces better results in broadcast-like simulations.\\
In the parallel case, the results are more subtle. The sensitivity to broadcast simulations is carried over in the parallel implementation, resulting in performance loss for both synchronization protocols. The optimistic case can still achieve good performance if the volume of inter-kernel messages is low. Both synchronization protocols are fast in non-cyclic simulations, whereas the conservative case is better suited whenever state-saving becomes expensive; the optimistic case, on the other hand, is better suited when a lookahead is not available.
Finally, the current optimistic implementation is ill-suited for memory constrained systems, especially in simulations where the frequency of reverts is high.
